---
title: "Project proposal"
author: "Team name: 7, El Puto's"
output: pdf_document
editor_options: 
---

```{r load-packages, message = FALSE,echo = FALSE,warning=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(ggplot2)
library(ggridges)
```

```{r setup, include = FALSE ,message=FALSE,warning=FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
```

## 1. Introduction

Research Question: Are there distinct differences between drug addicts
in the US who are submitted to rehab multiple times compared to drug
addicts who are submitted to rehab for the first time?

Our objective is to create a profile and identify similarities and
differences between individuals who have participated in rehabilitation
programs multiple times and those who are participating for the first
time. Our research aims to provide a better understanding of these
groups and match them with more appropriate treatments based on their
group profile and tendencies.

This analysis could make a significant contribution to the mental health
and drug addiction field. It can be challenging to match the right
treatment or facility with different addicts. Our vision is that if we
can create a distinct profile that affects the treatment given to the
individual, funds can be used more accurately. This could lead to the
creation of hospitals that are proficient with first-time patients and
others that are proficient with returning drug addicts.

This problem is significant because many recovered drug addicts are
constantly relapsing and are unable to leave the drug cycle. If we can
help in improving the treatment they receive, we may be able to help
them recover for good and live a better life.

Creating this profiling is difficult for many reasons, the first being
the credibility of the data, as many addicts might withhold information
on their admission to the rehab facility (for personal reasons,
skepticism in the process and in their anonymity etc.) Moreover, the
data is not a serial data, meaning we can't use prior admissions to
rehab for returning addict, a data segment that could have been helpful
to our cause.

Given that we have the admissions and discharges data from these
facilities, we wish to create a tree model that can classify our rows
based on the number of admissions to rehab a patient has, hopefully our
tree will catch accurately as many groups as possible. Our groups will
mainly focus on the number of times an addict was admitted to rehab, but
might also take into consideration the patient's preferable drug type,
and / or their mental health status upon admission. Than, we wish to
point out the most significant indicators for each group (class) using
the SHAP library


\pagebreak


## 2. Data

Our data set holds records of admissions to rehab facilities over the US
during the year of 2020. The data has both features from before
admission to the rehab facility, from the day of admission and of the
treatment offered to the patient. In our analysis we chose to focus on
data from before and from the day of admission as personal features that
might show differences in different groups in our research question.
Some of the original features were too specific to our purpose, so we
created 8 calculated features that use the original data and add to it
more suitable information. All the different features used for our work
are explained, both in meaning and in calculation method, and de-coded
in the README file attached as an appendix. The original data set in
encoded and all variables are numeric-labeled.


Quick explanation on the calculated features:

-   CALC_RACE - Categorical field for the RACE feature.
-   STATE_PARTY - The main party voted for in the patient's state
    presidential elections.
-   SUBSTANCE - categorical value, the type of most meaningful substance
    a patient is consuming (method in README). - ADDICTIVE_LEVEL - The
    highest addictive level (relative) for a substance consumed by the
    patient.
    -   FREQ - The highest frequency of drug consumption by the patient
        (self reported).
    -   FRSTUSE_LEGAL , FRSTUSE_ILLEGAL - The age of the first use in
        legal or illegal drug
    -   UNDER_INFLUENCE - Were drugs found in the patients body during
        admission?

After creating the calculated features and focus on the data features
relevant to us (in the README file) the data has been filtered. The
purpose of the filter is to ignore rows that might confuse our model.
Our project focuses on creating a relative ID for different groups (and
is not necessarily a prediction tool), so incomplete or ambiguous data
is ignored. We focused on a few issues (index is for identification in
the code):

1.  (1) Addicts that are addicted to 2 or more illegal drug types
    (uppers/ downers, over the counter) might compromise our profile,
    that takes into consideration the substance used. They are currently
    marked as SUNDAY_FUNDAY in SUBSTANCE column.

2.  (2) Null values (or unknown values) in original features:
    GENDER,MARSTAT, EDUC, EMPLOY, LIVARAG, PSOURCE, PSYPROB, NOPRIOR
    were removed.

3.  (3) Null values in calculated features: SUBSTANCE, FREQ, and in both
    FRSTUSE_LEGAL and FRSTUSE_ILLEGAL were removed.

Our final data set has 23 features (without caseID) and 579254 rows.

```{r ,warning=FALSE}
# Add original data import
admission_data <- read.csv("C:/Users/guy/Desktop/final_prop/data/TEDSA_PUF_2020.csv")
# /
#
# Feature engineering as described in the README and coded in the Appendix
#
# Filters and Selection to the engineered data
#
# Import new file
#/
admission_data <- read.csv("C:/Users/guy/Desktop/final_prop/data/admission_data_cooked.csv")

```

\pagebreak

## 3. Preliminary results

Exploratory data analysis, including some summary statistics and
visualizations, along with some explanation on how they help you learn
more about the problem. Obviously, you will add/implement more analysis
as you work on your final project, but for the proposal stage, we want
to see that project you're proposing is viable (and reasonable) to
accomplish within the Class's time frame.

On this section, our purpose is to show a basic analysis of our final
data set created above, as well as to show a few important relationships
in the data. First let's show statistical information on the final
features, and a glimpse into their first few values.

```{r}
# Get an upper view on the data
dim(admission_data)
summary(admission_data)

```

We are taking into consideration that our outcome might also include
separate profiles based on the calculated SUBSTANCE feature and the
original PSYPROB feature, as the patient's mental health and drug type
might also have a major influence on his/her profile. In order to
decide, correlation tests will be conducted between these 3 parameters.
The goal will be to see if there is a notable correlation between
SUBSTANCE \<-\> NOPRIOR and PSYPROB \<-\> NOPRIOR. If True, they will be
used as explanatory variables for our model. if False, they will have to
be represented in a different class and to be given with their own
profile.

\pagebreak


### First relationship: NOPRIOR \<-\> PSYPROB

As in the README, here is the decoded data labels:

0 - No priors

1 - 1 prior admission

2 - 2 prior admission

3 - 3 prior admission

4 - 4 prior admission

5 - 5 or more prior admission.

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.height=6, fig.width=10}
# Add original data import
admission_data <- read.csv("C:/Users/guy/Desktop/final_prop/data/admission_data_cooked.csv")
admissions <- admission_data[!is.na(admission_data$PSYPROB) & !is.na(admission_data$NOPRIOR),]

ggplot(admissions, aes(x = NOPRIOR, y = PSYPROB, color = factor(NOPRIOR))) +
  geom_point(stat = "summary", fun.y = "mean", size = 5) +
  xlab("NOPRIOR - Prior Admissions to Rehab") +
  ylab("PSYPROB - Psychiatric Assessment") +
  ggtitle("Psychiatric Assessment by Prior Admissions to Rehab") +
  scale_color_manual(values = c("#1A237E", "#303F9F", "#3949AB", "#9B59B6", "#BB8FCE", "#D7BDE2", "#E8DAEF")) +
  annotate("text", x = 4, y = 1.6, label = paste("Correlation = ", round(cor(admissions$NOPRIOR, admissions$PSYPROB), 2)), size = 5, color = "#8B0000")

```

It appears that there is a medium negative correlation between NOPRIOR
\<-\> PSYPROB so PSYPROB will be treated as an explanatory, X variable.

\pagebreak

### Second relationship: NOPRIOR \<-\> SUBSTANCE 

As in the README, here are the decoded data labels:

0 - legal in country

1 - over the counter prescriptions / other

2 - Stimulant

3 - Depressant

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=6, fig.width=10}
admissions <- admission_data[!is.na(admission_data$SUBSTANCE) & !is.na(admission_data$NOPRIOR),]

ggplot(admissions, aes(x = NOPRIOR, y = SUBSTANCE, color = factor(NOPRIOR))) +
  geom_point(stat = "summary", fun.y = "mean", size = 5) +
  xlab("NOPRIOR - Prior Admissions to Rehab") +
  ylab("SUBSTANCE - Consumed Substance Class") +
  ggtitle("Psychiatric Assessment by Consumed Substance Class") +
  scale_color_manual(values = c("#8B0000", "#AB2F40", "#E2A9A9", "#F0E68C", "#CDBE70", "#BDB76B")) +
  annotate("text", x = 1, y = 1.75, label = paste("Correlation = ", round(cor(admissions$NOPRIOR, admissions$SUBSTANCE), 2)), size = 5, color = "#1A237E")

```

It appears that there is a medium positive correlation between NOPRIOR
\<-\> SUBSTANCE so SUBSTANCE will be treated as an explanatory, X
variable.

next, we'll show how the data disperses in a few of our chosen
variables, to get a preliminary idea on the biggest most common features
and their values in the set.

### Categorical Features \<-\> NOPRIOR (Number of Prior Admissions to Rehab)

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=6, fig.width=10}
ggplot(admission_data, aes(x = NOPRIOR, y = as.factor(STATE_PARTY), fill = as.factor(STATE_PARTY), color = as.factor(STATE_PARTY))) + 
  geom_density_ridges(alpha = 0.5) + 
  ggtitle("Density of STATE_PARTY Category for NOPRIOR")

ggplot(admission_data, aes(x = NOPRIOR, y = as.factor(ADDICTIVE_LEVEL), fill = as.factor(ADDICTIVE_LEVEL), color = as.factor(ADDICTIVE_LEVEL))) + 
  geom_density_ridges(alpha = 0.5) + 
  ggtitle("Density of ADDICTIVE_LEVEL Category for NOPRIOR")

ggplot(admission_data, aes(x = NOPRIOR, y = as.factor(FREQ), fill = as.factor(FREQ), color = as.factor(FREQ))) + 
  geom_density_ridges(alpha = 0.5) + 
  ggtitle("Density of FREQ Category for NOPRIOR")

```

A few quick conclusions (remember, decoding of the Y axis's is in the
README file): - We see that in democrat states (2) there are much more
patients in 5th admission (and more), relative to the group size. One
might think that citizens in republican states (0) "give up" after the
first rehab. - We clearly see that the more addictive the consumed drug
is, the greater is the number of admissions to 5th rehab (and more) -
The frequency of use has a similar effect.

### First Use \<-\> Substance

```{r,echo=FALSE,fig.height=4, fig.width=10,warning=FALSE}

admission_data$GENDER <- as.numeric(admission_data$GENDER)
admission_data$PREG <- as.numeric(admission_data$PREG)
admission_data$VET <- as.numeric(admission_data$VET)
admission_data$ARRESTS <- as.numeric(admission_data$ARRESTS)
admission_data$SUBSTANCE <- as.character(admission_data$SUBSTANCE)

ggplot(admission_data, aes(y = FRSTUSE_LEGAL,
                  fill = SUBSTANCE)) +
  geom_bar(position = "fill") +
  labs(
    x = "Proportion",
    y = "FRSTUSE_LEGAL",
    fill = "SUBSTANCE",
    title = "Relation between Substance to first use of legal drug",
  ) + 
  ylim(-1,7)

ggplot(admission_data, aes(y = FRSTUSE_ILLEGAL,
                  fill = SUBSTANCE)) +
  geom_bar(position = "fill") +
  labs(
    x = "Proportion",
    y = "FRSTUSE_ILLEGAL",
    fill = "SUBSTANCE",
    title = "Relation between Substance to first use of illegal drug",
  ) + 
  ylim(-1,7)

admission_data$SUBSTANCE <- as.numeric(admission_data$SUBSTANCE)

```

By carefully examining this visualization we can see that the sooner a
person used a legal substance (age), the severity of the substance used
on the day of admission to rehab is usually higher, which might indicate
that the younger a person consume a legal drug for the first time, the
higher are the chances of this drug to turn into a gate-way drug in the
future (lead to use of worse drugs)

On the illegal drugs we see a clear and reversed conclusion: The older a
person is when consuming their first illegal drug, the worse is the drug
consumed on the day of admission to rehab.

\pagebreak

##### Categories disperse

A quick view on the relative quantities of each sub group out of the
total:

```{r,echo=FALSE,fig.height=6, fig.width=10,warning=FALSE}
# Set up a multi - paneled layout
par(mfrow = c(1,4))

# Create the first pie chart
gender_table <- table(admission_data$GENDER)
gender_labels <- c("Male", "Female")
names(gender_table) <- gender_labels
pie(gender_table, main = "Gender")

# Create the second pie chart
preg_table <- table(admission_data$PREG)
preg_labels <- c("Yes", "No")
names(preg_table) <- preg_labels
pie(preg_table, main = "Pregnent at Admission?")

# Create the third pie chart
vet_table <- table(admission_data$VET)
vet_labels <- c("Yes", "No")
names(vet_table) <- vet_labels
pie(vet_table, main = "Veteran?")

# Create the forth pie chart
ARRESTS_table <- table(admission_data$ARRESTS)
ARRESTS_labels <- c("None", "Once","Two or more times")
names(ARRESTS_table) <- ARRESTS_labels
pie(ARRESTS_table, main = "Arrests in past 30 days")
```


\pagebreak


## 4. Data analysis plan

What's next?

-   Our outcome (response, Y) is the feature NOPRIOR as we wish to
    define profiles for returning rehab patients.

-   Our predictor(explanatory, X) variables are, at least at the
    beginning (before proven not useful): AGE, GENDER, CALC_RACE,
    MARSTAT, EDUC, FRSTUSE_LEGAL, FRSTUSE_ILLEGAL, EMPLOY, PREG, VET,
    STATE_PARTY, LIVARAG, PRIMINC, ARRESTS, PSOURCE,
    DSMCRIT,FREQ_ATND_SELF_HELP, UNDER_INFLUENCE, ADDICTIVE_LEVEL,
    FREQ,PSYPROB, SUBSTANCE. These variables should compose a unique ID
    for each predicted class.

-   Method: We believe that the best model for our purpose will be a
    tree classifier model, so that every class in the outcome will have
    unique branches reaching to it. After defining a good enough tree,
    we'll use the SHAP library in order to catch the most important
    feature for each of our target classes. This way we'll be able to
    get unique ID's for returning addicts in comparison to first timers,
    based on their most important features.

-   To support our theory we'll need to prove that at least some of the
    classes we'll choose will have different characteristics (meaningful
    features in the decision tree). If we end up combining for example
    2nd and 3rd timers, that's okay, but we do expect to find
    differences at least between 1st timers in rehab and the rest.

-   Our work is divided to create load distribution, by each group
    member's availability. - Up until now one student was in charge of
    the calculated features, while the other 2 were in charge of the
    README file, method, proposal and learning the data. The exploratory
    data analysis was conducted by all 3 of us.

-   For the rest of the project we believe that the correlation and
    finalizing of the research question's target classes will be one
    student's job, another will be in charge of building the model, and
    the third will be in charge of analyzing the results. That being
    said, we'll all be active at least for QA in every step of the way.

Stay tuned for more!

\pagebreak

### Data README

```{r include_data_readme, comment=''}
cat(readLines('../data/README.md'), sep = '\n')
```



## Appendix

##### features engineering

In this Section We'll place our features engineering as described in the
data section.

```{r}
# Add original data import
admission_data <- read.csv("C:/Users/guy/Desktop/final_prop/data/TEDSA_PUF_2020.csv")

# Calculated features:

# First mutation of existing features (described in the README) - 
admission_data$PREG[admission_data$PREG == -9] <- 2
admission_data$VET[admission_data$VET == -9] <- 2
admission_data$ARRESTS[admission_data$ARRESTS == -9] <- 0
admission_data$RACE[admission_data$RACE == -9] <- 7
```

```{r}

# Create Party affiliation feature based on patients state
Republican <- c(1,2,5,18,20,21,22,28,29,30,31,38,40,45,46,47,48,49,54,56)
Democratic <- c(6,8,9,10,11,15,17,23,25,27,32,33,34,35,36,44,50)

# State party :
# Republican - 0
# Unstable / else - 1
# Democratic - 2

STATE_PARTY <- function(party){
    # If from a Democratic state
    if (party %in% Democratic) {
    return(2)
    # If If from a Republican state
  } else if (party %in% Republican) {
    return(0)
    # else - non-distinct-party affiliated state
    # If missing/unknown/not collected/invalid
  } else {
    return(1)
  }
}

tmp <- admission_data %>% select(STFIPS)

PARTY_v <- apply(tmp, 1, function(tmp) STATE_PARTY(tmp[1]))

# Activate calculated field

admission_data$STATE_PARTY <- PARTY_v


# Create new column FREQ containing the maximum value across FREQ1, FREQ2, and FREQ3
admission_data$FREQ <- apply(admission_data[, c("FREQ1", "FREQ2", "FREQ3")], 1, max, na.rm = TRUE)

# Create new feature UNDER_INFLUENCE on the day of admission
UNDER_INFLUENCE_v <- apply(admission_data[, c("ALCFLG", "COKEFLG", "MARFLG", "HERFLG", "METHFLG", "OPSYNFLG", "PCPFLG", "HALLFLG", "MTHAMFLG", "AMPHFLG", "STIMFLG", "BENZFLG", "TRNQFLG", "BARBFLG", "SEDHPFLG", "INHFLG", "OTCFLG", "OTHERFLG")], 1, function(admission_data) ifelse(sum(admission_data) > 0, 1, 0))
admission_data$UNDER_INFLUENCE <- UNDER_INFLUENCE_v
```

```{r}
# create a function SUB_ACCESS to assist in creating SUBSTANCE column
# legal -> TRUE , illegal -> FALSE
# vector of countries that marijuana is legal
weed_legal_states <- c(2,4,6,8,9,11,17,23,25,26,30,32,34,36,50,51)
SUB_ACCESS <- function(drug,state){
    # If alcohol
    if (drug == 2) {
    return(TRUE)
    # If weed and legal state
  } else if (drug == 4 & state %in% weed_legal_states) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

# Create a new function to indicate the type of the drug (by scale)
# legal in country - 0
# over the counter prescriptions / other - 1
# Stimulant - 2 
# Depressant - 3
# Irrelevant - -9 (if value is 1 (NONE) or -9 (unknown))

SUB_CATEGORY <- function(drug,state){
    # If Stimulant
    if (drug %in% c(3, 10, 11, 12)) {
    return(2)
    # If Depressant
  } else if (drug %in% c(5, 6, 7, 13, 14, 15, 16)) {
    return(3)
    #if value is 1 (NONE) or -9 (unknown)
  } else if (drug %in% c(1,-9)) {
    return(-9)
    # If legal in state
  } else if (SUB_ACCESS(drug,state) == TRUE) {
    return(0)
  # Over the counter / other
  } else {
    return(1)
  }
}

SUBSTANCE_CLASS <- function(drug1,drug2,drug3,state){
    # Return 'SUNDAY_FUNDAY' if there is a combination of active drugs (in README)
    # else: Return max value for all 3 substances in use
  c1 = SUB_CATEGORY(drug1,state)
  c2 = SUB_CATEGORY(drug2,state)
  c3 = SUB_CATEGORY(drug3,state)
  
  c_range <- c(-9, 0, 1, 2, 3)
  c_vector <- c(c1, c2, c3)
  c_unique <- unique(c_vector) # keeps only unique values in vector, without values that appears twice
  c_range_check <- c_unique %in% c_range[3:5] # take all unique values in range [1,3], holds TRUE and False
  
  if (sum(c_range_check) >= 2) {
    return('SUNDAY_FUNDAY')
  } else {
    return(max(c_vector))
  }
}

tmp <- admission_data %>% select(SUB1,SUB2,SUB3,STFIPS)

SUBSTANCE_v <- apply(tmp, 1, function(tmp) SUBSTANCE_CLASS(tmp[1],tmp[2],tmp[3],tmp[4]))
# Activate calculated field

admission_data$SUBSTANCE <- SUBSTANCE_v


```

```{r}
# Create vector for first use of legal drug
legal_drugs_firstuse = c()
illegal_drugs_firstuse = c()


FIRST_USE_LEGAL <- function(age1, drug1, age2, drug2, age3, drug3, state, legal_drugs_firstuse) {
  if (SUB_CATEGORY(drug1, state) == 0) {
    legal_drugs_firstuse <- c(legal_drugs_firstuse, age1)
  }
  if (SUB_CATEGORY(drug2, state) == 0) {
    legal_drugs_firstuse <- c(legal_drugs_firstuse, age2)
  }
  if (SUB_CATEGORY(drug3, state) == 0) {
    legal_drugs_firstuse <- c(legal_drugs_firstuse, age3)
  }
  if (length(legal_drugs_firstuse) == 0) {
    return (NA)
  } else {
    # Return min value for all 3 substances in use
    return (min(legal_drugs_firstuse))
  } 
}

FIRST_USE_ILLEGAL <- function(age1,drug1,age2,drug2,age3,drug3,state,illegal_drugs_firstuse){
  if (SUB_CATEGORY(drug1,state) > 0) {
    illegal_drugs_firstuse <- c(illegal_drugs_firstuse, age1)
  }
  if (SUB_CATEGORY(drug2,state) > 0) {
    illegal_drugs_firstuse <- c(illegal_drugs_firstuse, age2)
  }
  if (SUB_CATEGORY(drug3,state) > 0) {
    illegal_drugs_firstuse <- c(illegal_drugs_firstuse, age3)
  }
  if (length(illegal_drugs_firstuse) == 0) {
    return (NA)
  } else {
    # Return min value for all 3 substances in use
  return (min(illegal_drugs_firstuse))
  } 
}

tmp <- admission_data %>% select(FRSTUSE1,SUB1,FRSTUSE2,SUB2,FRSTUSE3,SUB3,STFIPS)

FRSTUSE_LEGAL_v <- apply(tmp, 1, function(tmp) FIRST_USE_LEGAL(tmp[1],tmp[2],tmp[3],tmp[4],tmp[5],tmp[6],tmp[7],legal_drugs_firstuse))
FRSTUSE_ILLEGAL_v <- apply(tmp, 1, function(tmp) FIRST_USE_ILLEGAL(tmp[1],tmp[2],tmp[3],tmp[4],tmp[5],tmp[6],tmp[7],illegal_drugs_firstuse))
# Activate calculated field

admission_data$FRSTUSE_LEGAL <- FRSTUSE_LEGAL_v
admission_data$FRSTUSE_ILLEGAL <- FRSTUSE_ILLEGAL_v

```

```{r}
# Define the severity of addiction potential for each drug category
# Non addictive / Unknown - 1 
# Low - 2
# High - 3
# Very high - 4

addiction_levels <- c("1" = 1,
                      "2" = 3,
                      "3" = 3,
                      "4" = 2,
                      "5" = 4,
                      "6" = 3,
                      "7" = 4,
                      "8" = 3,
                      "9" = 2,
                      "10" = 3,
                      "11" = 3,
                      "12" = 3,
                      "13" = 3,
                      "14" = 3,
                      "15" = 4,
                      "16" = 3,
                      "17" = 2,
                      "18" = 2,
                      "19" = 1,
                      "-9" = 1)
ADDICTION_LEVEL <- function(SUB1, SUB2, SUB3) {
  a1 = addiction_levels[as.character(SUB1)]
  a2 = addiction_levels[as.character(SUB2)]
  a3 = addiction_levels[as.character(SUB3)]
  max_value <- max(a1, a2, a3)
  return(max_value)
}

# Assign addictive level to the most addictive drug the patient funds
# create a new column for addiction potential



tmp <- admission_data %>% select(SUB1,SUB2,SUB3)

ADDICTIVE_LEVEL_v <- apply(tmp, 1, function(tmp) ADDICTION_LEVEL(tmp[1],tmp[2],tmp[3]))
# Activate calculated field

admission_data$ADDICTIVE_LEVEL <- ADDICTIVE_LEVEL_v
```

```{r}
CALC_RACE <- function(race){
    # If Black person
    if (race %in% c(4)) {
    return(1)
    # If White person
  } else if (race %in% c(5)) {
    return(2)
    # Other single race
  } else if (race %in% c(1,2,3,6,7,8,9)) {
    return(3)
    # If missing/unknown/not collected/invalid
  } else if (race %in% c(-9)) {
    return(-9)
  } 
}

tmp <- admission_data %>% select(RACE)

RACE_v <- apply(tmp, 1, function(tmp) CALC_RACE(tmp[1]))
# Activate calculated field

admission_data$CALC_RACE <- RACE_v

```

```{r}
# Take only relevant features for the final data set
admission_data <- admission_data %>% 
  select(CASEID, AGE, GENDER, CALC_RACE, MARSTAT, EDUC, FRSTUSE_LEGAL, FRSTUSE_ILLEGAL, EMPLOY, PREG, VET, STATE_PARTY, LIVARAG, PRIMINC, ARRESTS, PSOURCE, DSMCRIT,FREQ_ATND_SELF_HELP, UNDER_INFLUENCE, SUBSTANCE, ADDICTIVE_LEVEL, FREQ, PSYPROB, NOPRIOR) 
```

```{r}
# Apply filters to create the final data set

# FILTER (1)
admission_data <- admission_data[admission_data$SUBSTANCE != "SUNDAY_FUNDAY", ]
admission_data$SUBSTANCE <- as.integer(admission_data$SUBSTANCE)

#FILTER (2)
admission_data <- admission_data[admission_data$GENDER != -9 &
                                   admission_data$MARSTAT != -9 &
                                   admission_data$EDUC != -9 &
                                   admission_data$EMPLOY != -9 &
                                   admission_data$LIVARAG != -9 &
                                   admission_data$PSOURCE != -9 & admission_data$PSYPROB != -9 & 
                                   admission_data$NOPRIOR != -9, ]

#FILTER (3)
admission_data <- admission_data[!(is.na(admission_data$SUBSTANCE) | is.na(admission_data$FREQ) | 
                                     (is.na(admission_data$FRSTUSE_LEGAL) & is.na(admission_data$FRSTUSE_ILLEGAL))), ]

```

```{r}
# Write out the data to the repository to save re-calculations in the future
write.csv(admission_data,"C:/Users/guy/Desktop/final_prop/data/admission_data_cooked.csv",
          row.names = FALSE) 


```



### Source code

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
